### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@130106cead1e869da92e8484ce7935e60b6734ad 🚀

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.

@@

The Geographic Mosaic Theory of Coevolution (GMTC) attempts to
explicitly connect microevolutionary dynamics to the macroecology and
biogeography of symbiotic interactions.\textsuperscript{7} The GMTC
posits that coevolutionary processes among pairs\textsuperscript{8} or
complexes\textsuperscript{9} of species are structured in space by the
rippling effects of abiotic conditions onto evolutionary mechanism,
giving rise to fragmented systems with different structure and
ecologically dynamics over large spatial extents.\textsuperscript{10}
The GMTC predicts a spatial fragmentation of coevolutionary dynamics
under the joint action of three processes:\textsuperscript{11}
coevolutionary hot- and coldspots, which appear when the intensity of
\emph{interaction} (in terms of reciprocal fitness consequences) varies
spatially; selection mosaics, wherein the intensity of \emph{selection}
varies across space, driven by both the biotic complexity of the
community (locally diverse hosts and viruses are more biotically
complex) and the local favorability of the
environment;\textsuperscript{12} and trait remixing, which occurs when
coevolutionary dynamics are driven by by the arrival (or departure) of
\emph{functional traits}, through changes in community composition due
to invasions, meta-community dynamics, and disperal.

@@

Reassuringly, these predictions are validated by the sites where the
three highly pathogenic coronavirus are believed to have recently
emerged. While available information puts the spillover of SARS-CoV-2 in
a live animal market in Wuhan, China, the ultimate origin of the virus
is almost certainly in a divergent lineage of sarbecoviruses from the
Indochinese peninsula that was poorly characterized prior to the
pandemic.\textsuperscript{25--27} Similarly, the SARS-CoV outbreak began
in Guangdong province in 2002, reaching humans through small carnivore
bridge hosts, but was eventually traced back to a set of likely
progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
province;\textsuperscript{28} nearby, antibody evidence has indicated
human exposure to SARS-like viruses.\textsuperscript{29} MERS-CoV was
originally detected in Saudi Arabia, accompanied by a nearly identical
virus sequenced from an Egyptian tomb bat (\emph{Taphozous
perforatus})\textsuperscript{30}, but is widespread in camels in East
Africa and the Middle East, and may have reached its bridge host decades
earlier than originally supposed;\textsuperscript{31} as a result, the
geography of the original bat-to-camel transmission is still widely
regarded as uncertain. All of these are broadly consistent with the risk
factors we identify. Notably, India and west Africa are additional
hotspots that have yet to experience the emergence of a bat coronavirus
into human populations, but may still be at risk---particularly given
known gaps in bat surveillance,\textsuperscript{17} and a dense
population in both regions with global connectivity. In any of these
regions, surveillance on viral reservoirs can be paired with targeted
monitoring of high-risk human populations (i.e., those with regular
wildlife contact)\textsuperscript{32} for maximum impact.

@@

\leavevmode\hypertarget{ref-Olival2020PosZoo}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

@@

\leavevmode\hypertarget{ref-Rulli2021LanUse}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

@@

\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Worobey, M. \emph{et al.} The huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022).}

@@

\leavevmode\hypertarget{ref-Memish2013MidEas}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Memish, Z. A. \emph{et al.} Middle east respiratory
syndrome coronavirus in bats, saudi arabia. \emph{Emerging infectious
diseases} \textbf{19}, 1819 (2013).}

\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
neutralizing antibodies in camels, eastern africa, 1983--1997.
\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}

>>>>

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing SARS, MERS, and
COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.

@@

The Geographic Mosaic Theory of Coevolution (GMTC) attempts to
explicitly connect microevolutionary dynamics to the macroecology and
biogeography of symbiotic interactions.\textsuperscript{7} The GMTC
posits that coevolutionary processes among pairs\textsuperscript{8} or
complexes\textsuperscript{9} of species are structured in space by the
rippling effects of abiotic conditions onto evolutionary mechanisms,
giving rise to fragmented systems with different ecologies over large
spatial extents.\textsuperscript{10} The GMTC predicts a spatial
fragmentation of coevolutionary dynamics under the joint action of three
processes:\textsuperscript{11} coevolutionary hot- and coldspots, which
appear when the intensity of \emph{interaction} (in terms of reciprocal
fitness consequences) varies spatially; selection mosaics, wherein the
intensity of \emph{selection} varies across space, driven by both the
biotic complexity of the community (locally diverse hosts and viruses
are more biotically complex) and the local favorability of the
environment;\textsuperscript{12} and trait remixing, which occurs when
coevolutionary dynamics change when community-level \emph{functional
traits} change through meta-community dynamics.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from the Indochinese peninsula that was poorly
characterized prior to the pandemic.\textsuperscript{25--27} Similarly,
the SARS-CoV outbreak began in Guangdong province in 2002, reaching
humans through small carnivore bridge hosts, but was eventually traced
back to a set of likely progenitor viruses found in cave-dwelling
horseshoe bats in Yunnan province;\textsuperscript{28} nearby, antibody
evidence has indicated human exposure to SARS-like
viruses.\textsuperscript{29} MERS-CoV was originally detected in Saudi
Arabia, accompanied by a nearly identical virus sequenced from an
Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{30}, but
is widespread in camels in East Africa and the Middle East, and may have
reached its bridge host decades earlier than originally
supposed;\textsuperscript{31} as a result, the geography of the original
bat-to-camel transmission is still widely regarded as uncertain. All of
these are broadly consistent with the risk factors we identify. Notably,
India and west Africa are additional hotspots that have yet to
experience the emergence of a bat coronavirus into human populations,
but may still be at risk---particularly given known gaps in bat
surveillance,\textsuperscript{17} and a dense population in both regions
with global connectivity. In any of these regions, surveillance on viral
reservoirs can be paired with targeted monitoring of high-risk human
populations (i.e., those with regular wildlife
contact)\textsuperscript{32} for maximum impact.

@@

\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}

@@

\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
Land-use change and the livestock revolution increase the risk of
zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
Food} \textbf{2}, 409--416 (2021).}

@@

\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}

@@

\leavevmode\hypertarget{ref-Memish2013MidEas}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
syndrome coronavirus in bats, Saudi Arabia. \emph{Emerging infectious
diseases} \textbf{19}, 1819 (2013).}

\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
neutralizing antibodies in camels, Eastern Africa, 1983--1997.
\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}

### 2022-06-28 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@130106cead1e869da92e8484ce7935e60b6734ad 🚀

Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.

>>>>

Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing SARS, MERS, and
COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.
